---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# withdots

<!-- badges: start -->
<!-- badges: end -->

The `withdots()` function adds `...` to the argument list of a function if it does not already have it. This lets the function tolerate extraneous arguments that are passed to it.

## Installation

You can install the development version of withdots like so:

``` r
remotes::install_github("NikKrieger/withdots")
```

## Demo

The base R function `match()` has no `...` in its argument list:

```{r match}
match
```

Therefore, it can't handle extraneous arguments:

```{r error_match, error=TRUE}
match("z", letters, bad_arg = "error!")
```

But if we give it dots, then it can:

```{r wd}
library(withdots)

match_with_dots <- withdots(match)

match_with_dots("z", letters, can_now_handle = "junk arguments")
```

Functions that already have `...` in their argument list are returned as is:

```{r identical}
identical(lapply, withdots(lapply))
identical(c, withdots(c))
```

## Notes

### A note about primitive functions

If a function is a primitive function (see `?primitive`) with a well-defined argument list containing `...` (e.g., `c()`, `sum()`, `as.character()`), then `withdots()` will return it as is. The test for a "well-defined argument list," given a function `fn`, is `is.function(args(fn))`. 

#### primitive functions with well-defined argument lists not containing `...`

If the primitive has a well-defined argument list that does not contain `...` (e.g., `round()`, `is.na()`, `sqrt()`), then `withdots()` throws an error. To bypass this, all of these functions can be pre-processed with `rlang::as_closure()`, whose result can be then passed to `withdots()`. Observe:

```{r round}
# Observe that round() is a primitive function with no ... in its arg list:
round
```

```{r, rounderror, error=TRUE}
# So, we can't pass it to withdots() as is:
withdots(round)
```

```{r round_with_dots}
# But if we turn it into a closure, we can give it dots:
library(rlang)

round <- withdots(as_closure(round))

round
```

```{r round3}
# And now it can handle extraneous arguments:
round(45.78, digits = 1, junk, arguments)
```

**However**, keep in mind that the argument matching behavior of the result **_may_** be different from what is expected, since primitives **_may_** use nonstandard argument matching.

#### Primitive functions without well-defined argument lists

If the primitive function does not have a well-defined argument list (e.g., `[`, `~`, `function`, `for`), then `withdots()` throws an error. **Some** of these functions can be pre-processed with `rlang::as_closure()`, whose result definitely can be passed to `withdots()`. They are:

```{r can_be}
all_base <- getNamespaceExports("base")
all_base <- setNames(nm = all_base)
all_base <- lapply(all_base, function(x) getExportedValue("base", x))
all_primitives <- Filter(is.primitive, all_base)

primitive_non_well_defined_args <-
  Filter(function(fn) !is.function(args(fn)), all_primitives)

as_closure_coercible <-
  Filter(
    function(fn) tryCatch({as_closure(fn); TRUE}, error = function(e) FALSE),
    primitive_non_well_defined_args
  )

names(as_closure_coercible)
```

However, there are a handful of primitives that `rlang::as_closure()` is unwilling to process and are therefore ineligible for `withdots()`. They are:

```{r noncoercible}
as_closure_noncoercible <-
  Filter(
    function(fn) tryCatch({as_closure(fn); FALSE}, error = function(e) TRUE),
    primitive_non_well_defined_args
  )
  
names(as_closure_noncoercible)
```

### The `srcref` `attribute`. 

Many functions---including those created with `function()`---have a `srcref` `attribute`. When a function is `print`ed, `print.function()` relies on this `attribute` by default to depict the function's `formals` and `body`.

`withdots()` adds `...` via `formals<-`, which expressly drops `attributes` (see ``?`formals<-` ``). To prevent this loss, `withdots()` sets the function's `attributes` aside at the beginning and re-attaches them at the end. Normally, this would re-attach the original function's `srcref` `attribute` to the new function, making it so that the newly added `...` would not be depicted when the new function is `print`ed. For this reason, the old `srcref` `attribute` is dropped, and only the remaining `attributes` are re-attached to the new function.

Observe what would happen during `print`ing if **all** original `attributes` were naively added to the modified function:

```{r naive_withdots1}
# Create a function with no dots:
foo <- function(a = 1) {
  # Helpful comment
  a
}

# Give it important attributes that we can't afford to lose:
attr(foo, "important_attribute") <- "crucial information"
class(foo) <- "very_special_function"

# Print foo, which also prints its important attributes:
foo
```

```{r naive2}
# Save its attributes:
old_attributes <- attributes(foo)

# Add dots:
formals(foo)[["..."]] <- quote(expr = )

# See that the important attributes have been dropped:
foo
```

```{r naive3}
# Add the attributes back:
attributes(foo) <- old_attributes

# Print it again, and we see that the attributes have returned.
# However, the ... disappears from the argument list.
foo
```

```{r naive4}
# We know the actual function definitely has dots, since it can handle
# extraneous arguments:
foo(1, 2, junk, "arguments", NULL)
```

```{r naive5}
# Remove the "srcref" attribute, and the function is printed accurately.
# Furthermore, its important attributes are intact:
attr(foo, "srcref") <- NULL
foo

# Success! However, the comments in the body of the function are lost.
# Even so, this is better than inaccurate `print`ing.
```
