---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# withdots

<!-- badges: start -->
<!-- badges: end -->

The `withdots()` function adds `...` to the argument list of a function if it does not already have it. This lets the function tolerate extraneous arguments that are passed to it.

## Installation

You can install the development version of withdots like so:

``` r
remotes::install_github("NikKrieger/withdots")
```

## Demo

The base R function `match()` has no `...` in its argument list:

```{r match}
match
```

Therefore, it can't handle extraneous arguments:

```{r error_match, error=TRUE}
match("z", letters, bad_arg = "error!")
```

But if we give it dots, then it can:

```{r wd}
library(withdots)

match_with_dots <- withdots(match)

match_with_dots("z", letters, can_now_handle = "junk arguments")
```

Functions that already have `...` in their argument list are returned as is:

```{r lapply}
lapply
withdots(lapply)
```

```{r sum_wd}
c
withdots(c)
```

## Notes

### A note about `primitive` functions

If a function is primitive (see `?primitive`) and it has `...` in its argument list (e.g., `c()`, `sum()`, `as.character()`), it is returned as is. If the primitive function does not have `...` in its argument list, an error is thrown.

The user can bypass this by pre-processing the function with `rlang::as_closure()`. Observe:

```{r round}
# Observe that round() is a primitive function with no ... in its arg list:
round
```

```{r, rounderror, error=TRUE}
# So, we can't pass it to withdots() as is:
withdots(round)
```

```{r round_with_dots}
# But if we turn it into a closure, we can give it dots:
library(rlang)

round <- withdots(as_closure(round))

round
```

```{r round3}
# And now it can handle extraneous arguments:
round(45.78, digits = 1, junk, arguments)
```

**However**, keep in mind that the argument matching behavior of the result **_may_** be different from what is expected, since each primitive is special and **_may_** use nonstandard argument matching.

### The `srcref` `attribute`. 

Many functions---including those created with `function()`---have a `srcref` `attribute`. When a function is `print`ed, `print.function()` relies on this `attribute` by default to depict the function's `formals` and `body`.

`withdots()` adds `...` via `formals<-`, which expressly drops `attributes` (see ``?`formals<-` ``. To prevent this loss, `withdots()` sets the function's `attributes` aside at the beginning and re-attaches them to at the end. Normally, this would re-attach the original function's `srcref` `attribute` to the new function, making it so that the newly added `...` would not be depicted when the new function is `print`ed. For this reason, the old `srcref` `attribute` is dropped, and only the remaining `attributes` are re-attached to the new function.

Observe what would happen during `print`ing if **all** original `attributes` were naively added to the modified function:

```{r naive_withdots1}
# Create a function with no dots:
foo <- function(a = 1) {
  # Helpful comment
  a
}

# Give it important attributes that we can't afford to lose:
attr(foo, "important_attribute") <- "crucial information"
class(foo) <- "very_special_function"

# Print foo, which also prints its important attributes:
foo
```

```{r naive2}
# Save its attributes:
old_attributes <- attributes(foo)

# Add dots:
formals(foo)[["..."]] <- quote(expr = )

# See that the important attributes have been dropped:
foo
```

```{r naive3}
# Add the attributes back:
attributes(foo) <- old_attributes

# Print it again, and we see that the attributes have returned.
# However, the ... disappears from the argument list.
foo
```

```{r naive4}
# We know the actual function definitely has dots, since it can handle
# extraneous arguments:
foo(1, 2, junk, "arguments", NULL)
```

```{r naive5}
# Remove the "srcref" attribute, and the function is printed accurately.
# Furthermore, its important attributes are intact:
attr(foo, "srcref") <- NULL
foo
```

Success! However, the comments in the `body()` of the function are lost. Even so, this is better than inaccurate `print`ing.

